---
title: "02-R_Analysis"
output: html_document
date: "2023-10-09"
---

# R Analysis

## Pre-requisites

Required Libraries:
```{r}
if(!require("ggplot2"))install.packages("ggplot2")

library(ggplot2)
library(corrplot)
library(caret)
library(rpart)
library(reshape2)  # The reshape2 package is used for data reshaping.
library(dplyr) # The dplyr package is used for data manipulation.
library(rpart.plot)
```

## Data

Load the data:
```{r}
data <- read.table(path_wd("..","data","raw","smoking_driking_dataset.csv"),header=TRUE, sep = ",")
data <- as.data.frame(data)
df <- as.data.frame(data)
data$DRK_YN <- ifelse(data$DRK_YN == "Y", 1, 0)       # MUST ONLY BE RUN ONCE !
data$sex <- ifelse(data$sex == "Male", 1, 0)
```


```{r}
typeof(data)
class(data)
dim(data)
head(data)
summary(data)
str(data)
```
Remove outliers e.g. waistline 999
```{r}

```


Below is an explanation of these variables, which represent body data:

Sex	male, female	
age	round up to 5 years	
height	round up to 5 cm[cm]	
weight	[kg]	
sight_left	eyesight(left)	
sight_right	eyesight(right)	
hear_left	hearing left, 1(normal), 2(abnormal)	
hear_right	hearing right, 1(normal), 2(abnormal)	
SBP	Systolic blood pressure[mmHg]	 
DBP	Diastolic blood pressure[mmHg]	 
BLDS	BLDS or FSG(fasting blood glucose)[mg/dL]	
tot_chole	total cholesterol[mg/dL]	 
HDL_chole	HDL cholesterol[mg/dL]	HDL 
LDL_chole	LDL cholesterol[mg/dL]	LDL 
triglyceride	triglyceride[mg/dL]	
hemoglobin	hemoglobin[g/dL]	
urine_protein	protein in urine, 1(-), 2(+/-), 3(+1), 4(+2), 5(+3), 6(+4)	 
serum_creatinine	serum(blood) creatinine[mg/dL] 
SGOT_AST	SGOT(Glutamate-oxaloacetate transaminase) AST(Aspartate transaminase)[IU/L]	 AST
SGOT_ALT	ALT(Alanine transaminase)[IU/L]	 ALT
gamma_GTP	y-glutamyl transpeptidase[IU/L]	
SMK_stat_type_cd	Smoking state, 1(never), 2(used to smoke but quit), 3(still smoke)	
DRK_YN	Drinker or Not


We first convert the drinking data into numbers to facilitate our binary classification
```{r}
# Convert DRK_YN and sex columns to numeric encoding
df$DRK_YN <- as.numeric(factor(df$DRK_YN))
df$sex <- as.numeric(factor(df$sex))
```


```{r}
colSums(is.na(data))
```

### Test/Train split

```{r}
# Assuming 'dataset' is your data frame
predictors <- data[, -which(names(data) == "DRK_YN")]
target <- data$DRK_YN

# Split the data into training and testing sets
set.seed(0)  # Set a random seed for reproducibility
splitIndex <- createDataPartition(target, p = 0.8, list = FALSE)
X_train <- predictors[splitIndex, ]
X_test <- predictors[-splitIndex, ]
Y_train <- target[splitIndex]
Y_test <- target[-splitIndex]
```


## EDA and plots

```{r}
# Sample 10,000 rows from your_data
sada <- data[sample(nrow(data), 10000), ]
targetsada <- sada$DRK_YN

# 'sada' now contains 10,000 randomly selected rows from 'data'
head(sada)
```


```{r}
# Create a scatter plot
plot(sada$weight, sada$height, 
     xlab = "weight", ylab = "height", 
     main = "Scatter Plot of x vs. y")
```

```{r}
plot(sada$waistline,sada$gamma_GTP)
```

```{r}
max(data$waistline)
summary(data$waistline)
```

```{r}
ggplot(sada, aes(x = `hemoglobin`, group = targetsada, fill = targetsada)) +
  geom_density(alpha = 0.5)
```

##  Feature Selection and Correlation - Test/train split + Feature selection + correlation/PCA

```{r}
correlation_matrix <- cor(data)

# Get the absolute correlation values of "target" with other variables
target_correlations <- abs(correlation_matrix[,"DRK_YN"])

# Sort the correlations in descending order
sorted_correlations <- sort(target_correlations, decreasing = TRUE)

# Print the sorted correlations
print(sorted_correlations)
```

```{r}
corrplot(correlation_matrix, method = "color")
```

If we count the variables that have the highest impact on drinking, we find that whether drinking has a large positive correlation with the four variables of sex, height, smoking, and hemoglobin concentration, and a large negative correlation with age. 
Among them, height has the greatest impact, which may be unexpected.

One possible explanation is that there is a large positive correlation between body size and gender, which is caused by the fact that men are generally taller than women. All we have to do is look at the relationship between gender and height, hemoglobin and smoking

Visualizing the correlation matrix plot allows us to see this very quickly.No doubt, height, weight, hemoglobin and smoking have a very large positive correlation with gender!
This means that many of these variables work together.

### picking only uncorrelated features

```{r}
# Assuming 'correlation_matrix' is your correlation matrix
#high_correlation_threshold <- 0.85  # Define a correlation threshold

# Find highly correlated features
#highly_correlated <- findCorrelation(correlation_matrix, cutoff = high_correlation_threshold)

# Remove highly correlated features
#reduced_features <- X_train[, -highly_correlated]
#print(reduced_features)
```

### PCA

```{r}
# Scale the data for PCA (optional but recommended)
scaled_data <- scale(X_train)

# Apply PCA
pca_result <- prcomp(scaled_data, scale. = TRUE)

# Explained variance by each principal component
summary(pca_result)

# Retrieve transformed data with reduced dimensions
reduced_data <- as.data.frame(predict(pca_result))
```

We take the first 12 principal components which capture 80% of the variance

```{r}
reddata <- reduced_data[,1:12]
correlation_matrix2 <- cor(reddata)
corrplot(correlation_matrix2, method = "color")
```

We'll reduce the PCA result data frame into our 12 PCs that capture our vairance tolerance

```{r}
data_pca <- as.data.frame(pca_result$x[, 1:12])
head(data_pca)
```

We now create training and test data based on PC1-PC12 to see how it compares ahead with model accuracy

```{r}
# Assuming 'dataset' is your data frame
predictors_pca <- data_pca[, -which(names(data) == "DRK_YN")]
target_pca <- data$DRK_YN

# Split the data into training and testing sets
set.seed(0)  # Set a random seed for reproducibility
splitIndex <- createDataPartition(target, p = 0.8, list = FALSE)
X_train_pca <- predictors_pca[splitIndex, ]
X_test_pca <- predictors_pca[-splitIndex, ]
Y_train_pca <- target_pca[splitIndex]
Y_test_pca <- target_pca[-splitIndex]
```


## XGBoost

XGBoost, short for eXtreme Gradient Boosting, It falls under the category of ensemble learning methods and is based on decision tree models. XGBoost trains models in a gradient-boosting manner.
 It falls under the category of ensemble learning methods and is based on decision tree models. XGBoost trains models in a gradient-boosting manner. 
code based on (https://www.kaggle.com/code/raman209/prediction-of-drinkers-using-body-signals)

```{r}
# Select the feature columns
X <- df[, !names(df) %in% c("DRK_YN")]
# Select the target variable column
y <- df$DRK_YN
```

Since we mentioned before that drinking is related to age, let’s try to visualize the age distribution

```{r}
# Create a ggplot chart
p <- ggplot(data = df, aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "lightblue", color = "black") +
  geom_density(color = "red") +
  labs(x = "Age", y = "Frequency", title = "Distribution of Age")

# Display the chart
print(p)
```
The age data are mainly concentrated in the 35-65 age group, which shows that the subjects who provided the data are generally middle-aged and elderly people.

We still split the data into training data and test data according to the ratio of 80%
```{r}
# Set a random seed to ensure reproducible results
set.seed(123)

test_size <- 0.2  # 20% of the data is used for testing

# Split the data using the sample.split function
split <- sample.split(y, SplitRatio = 1 - test_size)

# Create the training set
X_train <- X[split, ]
y_train <- y[split]

# Create the test set
X_test <- X[!split, ]
y_test <- y[!split]
```

Now we use xgboost for data fitting, and the number of boost rounds is set to 100.This number was obtained later through subsequent attempts.

```{r}
# Install the xgboost package if not already installed
if (!require("xgboost")) install.packages("xgboost")
library(xgboost)

# Create an XGBoost classifier model round = 100
model <- xgboost(data = as.matrix(X_train), label = y_train, objective = "binary:logistic", nrounds = 100)

# Use the model for predictions
y_pred <- predict(model, as.matrix(X_test))
# Print the prediction results
print(y_pred)
```

Next we test the accuracy of the model using a confusion matrix.

```{r}
# Custom rounding function
custom_round <- function(x) {
  ifelse(x < 0.5, 0, 1)
}

# Convert y_pred to binary classification labels
y_pred_binary <- ifelse(y_pred >= 0.5, 1, 0)

# Convert y_test to a factor
y_test <- factor(y_test, levels = c(0, 1))

# Convert y_pred_binary to a factor
y_pred_binary <- factor(y_pred_binary, levels = c(0, 1))

# Ensure they have the correct levels
levels_y_test <- levels(factor(y_test))
y_pred_binary <- factor(y_pred_binary, levels = levels_y_test)

# Now you can calculate accuracy and generate a classification report
accuracy <- confusionMatrix(data = y_pred_binary, reference = y_test)$overall["Accuracy"]
cat("Accuracy: ", format(accuracy, nsmall = 2), "\n")

# Generate a classification report
confusion_matrix <- confusionMatrix(data = y_pred_binary, reference = y_test)
print(confusion_matrix)
```

And the recall rate is:
```{r}
# Calculate True Positives (TP) and False Negatives (FN)
TP <- sum(y_test == 1 & y_pred_binary == 1)
FN <- sum(y_test == 1 & y_pred_binary == 0)

# Calculate Recall
recall <- TP / (TP + FN)

cat("Recall (True Positive Rate): ", format(recall, nsmall = 2), "\n")

```

xgboost did not significantly improve the accuracy of prediction, which may be due to many reasons.

## Decision Trees

```{r}
# Assuming X_train is the matrix/data.frame of features and Y_train is the target variable
#my_tree <- rpart(Y_train ~ ., data = cbind(X_train, Y_train), method = "class")
tree <- rpart(Y_train ~ ., data = X_train, method = "class")
tree_pca <- rpart(Y_train_pca ~ ., data = X_train_pca, method = "class")
```

```{r}
test_pred <- predict(tree, X_test, type = "class")
test_pred_pca <- predict(tree_pca, X_test_pca, type = "class")

confusionMatrix(test_pred, as.factor(Y_test))
confusionMatrix(test_pred_pca, as.factor(Y_test_pca)[1:38412])
```



