---
title: "02-R_Analysis"
output: html_document
date: "2023-10-09"
---

# R Analysis

## Pre-requisites

Required Libraries:
```{r}
# for *Yifan's part*
if(!require("ggplot2"))install.packages("ggplot2")
if(!require("ggpubr"))install.packages("ggpubr")
if(!require("gridExtra"))install.packages("gridExtra")

library(gridExtra)
library(ggpubr)
library(ggplot2)

# For manipulating data
library(magrittr)
library(rprojroot)
library(forcats)
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
library(readr)

# For descriptive statistics and graphing
library(skimr)
library(scales)
library(gridExtra)
```

Load the data:
```{r}
selected_file <- file.choose()
italy <- read.csv(selected_file)
```

## Basic Exploratory Data Analysis

For some EDA we first require the date to have a date object

Create a Date object data frame:
```{r}
italydf <- read_csv(selected_file, 
                    col_types = cols(Time = col_date(format = "%m/%d/%Y"))) %>% 
  setNames(c("Time", "AT")) %>% 
  arrange(Time)
```

```{r}
typeof(italydf)
```


The 'summary()' fuction is a good place to start however, we may alternately use 'skimr::skim()'

```{r}
italydf %>% skim()
```
It is important however, to not only use statistical summaries but also plot the data. We'll start with a simple line chart.

```{r}
italydf %>% 
  ggplot(aes(x = Time, y = AT)) +
  geom_line()

print(italydf)
```





